---
title: "MP4"
author: "Jonah Kotzen, Ava Skogstrom, Conor Booth"
date: "2024-04-17"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

```

```{r read average, include=FALSE}
alcidae <- read.table("Booth_Kotzen_Skogstrom_alcid.txt", sep = ",", header = TRUE)
laridae <- read.table("Booth_Kotzen_Skogstrom_larid.txt", sep = ",", header = TRUE)

```

# Total Wing Area (A)
We used the ruler and shape info features in SketchUp to determine A
A = chord (c) *x* span (s) (an approximation)
Laridae: c = 20.6623 mm, s = 60.1978 mm, A = 1243.825 sq mm
Alcidae: c = 42 mm, s = 52 mm, A = 2184 sq mm

# Span (S)
We used the ruler and shape info features in SketchUp to determine S
Laridae: 60.1978 mm
Alcidae: 52 mm

# Wetted Area (Aw)
We used the ruler and shape info features in SketchUp to determine Aw
The wetted area is surface area of the wing in contact with the air flow
Laridae: 165.368118 sq mm
Alcidae: 104 sq mm

# Comparison of Lift over Time b/w Specimens

```{r graph 1, include=TRUE}

ggplot() +
  geom_line(data = alcidae, aes(x = s, y = lift, color = "Alcidae Lift"), size = 1) +
  geom_line(data = laridae, aes(x = s, y = lift, color = "Laridae Lift"), size = 1) +
  labs(title = "Comparison of Lift over Time", x = "Time (s)", y = "Lift Force (N)") +
  scale_color_manual(values = c("Alcidae Lift" = "blue", "Laridae Lift" = "green")) +
  theme_minimal()

```

# Comparison of Drag over Time b/w Specimens

```{r graph 2, include=TRUE}

ggplot() +
  geom_line(data = alcidae, aes(x = s, y = drag, color = "Alcidae Drag"), size = 1) +
  geom_line(data = laridae, aes(x = s, y = drag, color = "Laridae Drag"), size = 1, linetype = "dashed") +
  labs(title = "Comparison of Drag over Time", x = "Time (s)", y = "Drag Force (N)") +
  scale_color_manual(values = c("Alcidae Drag" = "red", "Laridae Drag" = "purple")) +
  theme_minimal()
```

# Calculations

```{r calculations1, include=FALSE}
# Calculate averages for Alcidae
average_lift_alcidae <- mean(alcidae$lift)
average_drag_alcidae <- mean(alcidae$drag)

# Calculate averages for Laridae
average_lift_laridae <- mean(laridae$lift)
average_drag_laridae <- mean(laridae$drag)

# Create a tibble with average lift and drag for both bird families
averages_tibble <- tibble(
  Bird_Family = c("Alcidae", "Laridae"),
  Average_Lift = c(average_lift_alcidae, average_lift_laridae),
  Average_Drag = c(average_drag_alcidae, average_drag_laridae)
)
```

```{r calc2, include=TRUE}
averages_tibble
```

# Interpretations

# References
